### 2017-11-13 Draft2/Appending.Dynamic.tex
Work
Finish more

I now proceed to asymptotically bound $\FuncWrites(n)$:
\begin{align*}
\FuncWrites(n) &= n + \left( \frac{\VarGrowthFactor^\VarUseful - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity\\
n + \left( \frac{n / \VarInitCapacity - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \FluteLeq \FuncWrites(n) &\FluteLess n + \left( \frac{\VarGrowthFactor{n} / \VarInitCapacity - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n - \left( \frac{\VarInitCapacity}{\VarGrowthFactor - 1} \right) \FluteLeq \FuncWrites(n) &\FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n - \left( \frac{\VarInitCapacity}{\VarGrowthFactor - 1} \right)\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \FluteLeq \FuncWrites(n) &\FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n
\end{align*}
\HdrSpaceComplex

I wish to find the space allocated when $n$ items are appended to a dynamic array. I call this quantity the \textbf{space cost}, denote it $\FuncSpace(n)$, and define it as the total length of buffers allocated by $n$ $\FuncAppend$ calls. Now, I derive a formula for $\FuncSpace(n)$.

First, from the definition of $\VarList.\FieldCapacity$, note that a dynamic array's capacity is the length of the buffer it stores its items in. Then a buffer of length $c$ is allocated at some point if and only if $c \in \VarCapacitySeq$. Then the total length of those buffers is
\begin{align*}
\FuncSpace(n) &= \sum_i {\VarCapacitySeq_i}\\
&= \VarInitCapacity + \VarGrowthFactor\VarInitCapacity + \VarGrowthFactor^2\VarInitCapacity + \ldots + \VarGrowthFactor^\VarUseful\VarInitCapacity\\
&= \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
Using Lemma \ref{lem:ToVarUsefulPowerBounds} again, I asymptotically bound $\FuncSpace(n)$:
\begin{align*}
\left( \frac{\VarGrowthFactor{n} / \VarInitCapacity - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \FluteLeq \FuncSpace(n) &\FluteLess \left( \frac{\VarGrowthFactor^2{n} / \VarInitCapacity - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n - \left( \frac{\VarInitCapacity}{\VarGrowthFactor - 1} \right) \FluteLeq \FuncSpace(n) &\FluteLess \left( \frac{\VarGrowthFactor^2}{\VarGrowthFactor - 1} \right) n - \left( \frac{\VarInitCapacity}{\VarGrowthFactor - 1} \right)\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \FluteLeq \FuncSpace(n) &\FluteLess \left( \frac{\VarGrowthFactor^2}{\VarGrowthFactor - 1} \right) n
\end{align*}

>>>>

I can build on this inequality to receive asymptotic bounds for $\FuncWrites(n)$:
\begin{align*}
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful - 1 \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac n {\VarGrowthFactor - 1} &\FluteLeq \left( \frac{\VarGrowthFactor^\VarUseful - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \FluteLess \frac {\VarGrowthFactor{n}} {\VarGrowthFactor - 1}\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n &\FluteLeq \FuncWrites(n) \FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n
\end{align*}
\HdrSpaceComplex

I wish to find the space allocated when $n$ items are appended to a collection. I call this quantity the \textbf{space cost}, denote it $\FuncSpace(n)$, and define it as the total length of buffers allocated by $n$ appends. Now, I derive a formula for $\FuncSpace(n)$.

First, from the definition of $\VarList.\FieldCapacity$, note that a dynamic array's capacity is the length of the buffer it stores its items in. Then a buffer of length $c$ is allocated at some point if and only if $c \in \VarCapacitySeq$. Then the total length of those buffers is
\begin{align*}
\FuncSpace(n) &= \sum_i {\VarCapacitySeq_i}\\
&= \VarInitCapacity + \VarGrowthFactor\VarInitCapacity + \VarGrowthFactor^2\VarInitCapacity + \ldots + \VarGrowthFactor^\VarUseful\VarInitCapacity\\
&= \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
Using Lemma \ref{lem:ToVarUsefulPowerBounds} again, I asymptotically bound $\FuncSpace(n)$:
\begin{align*}
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac {\VarGrowthFactor n} \VarInitCapacity &\FluteLeq \VarGrowthFactor^{\VarUseful + 1} \FluteLess \frac {\VarGrowthFactor^2 n} \VarInitCapacity\\
\frac {\VarGrowthFactor n} \VarInitCapacity &\FluteLeq \VarGrowthFactor^{\VarUseful + 1} - 1 \FluteLess \frac {\VarGrowthFactor^2 n} \VarInitCapacity\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n &\FluteLeq \FuncSpace(n) \FluteLess \left( \frac{\VarGrowthFactor^2}{\VarGrowthFactor - 1} \right) n
\end{align*}
